---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# datacitation  [![Build Status](https://travis-ci.org/benmarwick/datacitation.svg?branch=master)](https://travis-ci.org/benmarwick/datacitation) 

This repository contains the data and code for our paper:

> Marwick, B., and S. Pilaar Birch (2018). "A standard for the scholarly citation of archaeological data as an incentive to data sharing" _Advances in Archaeological Practice_

Our pre-print is online here: 

> Marwick, B., & S. Pilaar Birch (2018). A Standard for the Scholarly Citation of Archaeological Data as an Incentive to Data Sharing. Retrieved from <https://osf.io/preprints/socarxiv/py4hz>

### Citation

Please cite this compendium as:

> Marwick, B., & S. Pilaar Birch, (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for the article "A standard for the scholarly citation of archaeological data as an incentive to data sharing"_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/10.17605/OSF.IO/KSRUZ>

### Installation

You can install datacitation from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("benmarwick/datacitation")
```

### Licenses

**Text and figures:**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code:** See the [DESCRIPTION](DESCRIPTION) file

**Data**: [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse

### Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

<sup><sub>This compendium was created with [rrtools](https://github.com/benmarwick/rrtools).</sup></sub>
